### 2020-06-18 report/sections/evalconcl/eval.tex
Report: feedback on benchmarks

In this chapter, we
evaluate the expressiveness
and performance of our generated APIs.

%TODO: add when finalised
%with respect to the main objective
%of \dots

@@

>>>>

In this chapter, we
evaluate the expressiveness
and performance of our generated APIs
with respect to the main objective of
our project in
providing
developers with a development workflow
that offers communication safety guarantees in modern
web programming through multiparty session types.

@@

\section{Summary}
We have demonstrated
the expressiveness of our generated APIs
through implementing two web services describing
multiparty sessions. We highlighted how our
approach towards session-typed GUI programming
prevents channel reuse and 
makes it intuitive to implement the game board
of \tprotocol{Noughts and Crosses}.
We also showed
that the extensions added to \nodecodegen
and \reactcodegen
to support \newtheory
are transparent to the developer.

Through performance benchmarks against a
baseline implementation of the \tprotocol{Ping Pong}
protocol, we analysed the overhead of our implementation
and reasoned that this overhead will minimise when compared
with baseline implementations of more complex, multiparty
protocols involving routed communication.

### 2020-06-18 report/sections/evalconcl/eval/benchmarks.tex
Report: feedback on benchmarks

Whilst web applications implementing our generated APIs enjoy
communication safety guarantees, the presence of the session runtime acts
as an additional layer of abstraction between the application logic and the
WebSocket transport, which is likely to present a performance trade-off.

@@

The interested reader may follow the instructions under
\texttt{benchmarks/README.md} in the project repository \cite{repo}
to run the benchmarks and visualise the logs using the interactive
notebook in the same directory.

We run the experiments under a network of latency 0.165ms
(64 bytes ping), and repeat each experiment 20 times.
Execution time measurements  are taken using a machine 
equipped with Intel i7-4850HQ CPU (2.3 GHz, 4 cores, 8 threads), 
16 GB RAM, macOS operating system version 10.15.4, 
Node.js runtime version 12.12.0, and
TypeScript compiler version 3.7.4.
We standardise all packages used in the front- and back-end
implementations across experiments. Details can be found in their
corresponding \texttt{package.json} manifests 
(\cref{section:evalbenchmarkpackagejson}).

@@

\subsection{Overhead}
We compare the total execution time (\textit{Exec. Time}) 
and execution time per round trip (\textit{Exec. Time / Ping-Pong}) -- 
averaged over 20 repeated experiments -- across the three implementations,
for $n \in \{10^2, 10^3, 10^4\}$.
We summarise the results in \cref{table:overhead}.

@@

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
socket.onmessage = ({ data }) => {
	const { label, payload } = JSON.parse(data.toString());
	if (label === 'PING') {
		let count: number = payload[0];
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			socket.send(JSON.stringify({
				label: 'BYE',
				payload: [count],
			}));
		} else {
			socket.send(JSON.stringify({
				label: 'PONG',
				payload: [count],
			}));
		}
	} else {
		throw new Error(`Unrecognised label: ${label}`);
	}
}
\end{lstlisting}
\caption{\texttt{bare} implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
// NOTE: `Implementation' namespace omitted for brevity.

const logic = new S14({
	PING: (count) => {
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			return new S16([Labels.S16.BYE, [count], new S15()]);
		} else {
			return new S16([Labels.S16.PONG, [count], logic]);
		}
 	}
});
\end{lstlisting}
\caption{\texttt{mpst} implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongmpst}
\end{subfigure}
\caption{Comparison of \texttt{bare} and \texttt{mpst}
implementations for \tprotocol{Ping Pong} \trole{Svr}}
\label{fig:nodepingpong}
\end{figure}

>>>>

Whilst web applications implementing 
our generated APIs enjoy
communication safety guarantees, 
the presence of the session runtime acts
as an additional layer of abstraction between 
the application logic and the
WebSocket transport, which presents
a performance trade-off; we observe an
increase in the time taken to perform
each channel action, as a result of
the session runtime intercepting channel actions
and performing additional logic to ensure
that the endpoint conforms to the protocol.

@@

Code is available at
\url{https://github.com/ansonmiu0214/SessionTS-Benchmarks}.
Interested readers may follow the \texttt{README.md}
to run the benchmarks and visualise the logs using the 
interactive notebook in the same directory.

We run the experiments under a network of latency 0.165ms
(64 bytes ping), and repeat each experiment 20 times.
Execution time measurements  are taken using a machine 
equipped with Intel i7-4850HQ CPU (2.3 GHz, 4 cores, 8 threads), 
16 GB RAM, macOS operating system version 10.15.4, 
Node.js runtime version 12.12.0, and
TypeScript compiler version 3.7.4.
We standardise all packages used in the front- and back-end
implementations across experiments. Details can be found in
\cref{section:evalbenchmarkpackagejson}.

@@

\subsection{Overhead}
The execution time is the total time taken
for the \trole{Client} to connect to the \trole{Server}
and complete the parameterised number of round trips.
We compare the total execution time (\textit{Exec. Time}) 
and execution time per round trip (\textit{Exec. Time / Ping-Pong}) -- 
averaged over 20 repeated experiments -- across the three implementations,
for $n \in \{10^2, 10^3, 10^4\}$.
We summarise the results in \cref{table:overhead}.

@@

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
socket.onmessage = ({ data }) => {
	const { label, payload } = JSON.parse(data.toString());
	if (label === 'PING') {
		let count: number = payload[0];
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			socket.send(JSON.stringify({
				label: 'BYE',
				payload: [count],
			}));
		} else {
			socket.send(JSON.stringify({
				label: 'PONG',
				payload: [count],
			}));
		}
	} else { throw new Error(`Unrecognised label: ${label}`); }
}
\end{lstlisting}
\caption{\texttt{bare} implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
const logic = new Implementat.S14({
	PING: (count) => {
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			return new Implementation.S16([
				Labels.S16.BYE, [count], new Implementation.S15()
			]);
		} else {
			return new Implementation.S16([
				Labels.S16.PONG, [count], logic
			]);
		}
 	}
});
\end{lstlisting}
\caption{\texttt{mpst} implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongmpst}
\end{subfigure}
\caption{Comparison of \texttt{bare} and \texttt{mpst}
implementations for \tprotocol{Ping Pong} \trole{Svr}}
\label{fig:nodepingpong}
\end{figure}

### 2020-06-18 report/sections/evalconcl/eval/twobuyer.tex
Report: feedback on benchmarks

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
new Implementation.Initial({
	[Labels.S32.title]: (title) => {
		return new Implementation.S34([
			Labels.S34.quote, [100], new Implementation.S35([
				Labels.S35.quote, [100], new Implementation.S36({
					[Labels.S36.buy]: () => {
						return new Implementation.Terminal();
					},
					[Labels.S36.cancel]: () => {
						return new Implementation.Terminal();
					}
				})
			])
		])
	}
})
\end{lstlisting}
\captionof{lstlisting}{Two Buyer Seller Implementation}
\label{lst:evaltwobuyerseller}
\end{figure}

>>>>

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
new Implementation.Initial({
	[Labels.S32.title]: async (title) => {
		const quote = await db.getQuote(title);
		return new Implementation.S34([
			Labels.S34.quote, [quote], new Implementation.S35([
				Labels.S35.quote, [quote], new Implementation.S36({
					[Labels.S36.buy]: () => {
						return new Implementation.Terminal();
					},
					[Labels.S36.cancel]: () => {
						return new Implementation.Terminal();
					}
				})
			])
		])
	}
})
\end{lstlisting}
\captionof{lstlisting}{Two Buyer Seller Implementation}
\label{lst:evaltwobuyerseller}
\end{figure}
